Explore_And_Summarize_Data_RedWine by Shipra_Gupta
========================================================



```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(ggthemes)
library(gridExtra)
library(scales)
library(CARS)
library(memisc)
library(lattice)
library(MASS)
library(GGally)
library(RColorBrewer)
library(reshape)
library(knitr)

getwd()

```

```{r echo=FALSE, Load_the_Data}

# Load the Data
list.files()
wq<-read.csv("wineQualityReds.csv")


```

I am going to explore the Quality of red wines 


# Univariate Plots Section

```{r echo=FALSE}

qplot(x=quality,data=wq,binwidth=.1)

```

As we can see above most of the wines fall in quality range 5 to 6 .

```{r echo=FALSE, Univariate_Plots}


qplot(x=alcohol,data=wq,binwidth=.1)+
  scale_x_continuous(breaks = c(1:15))

```

Above we can see the alcohol content of 9.6 shows maximum counts of red wine.



```{r echo=FALSE}

qplot(x=log10(residual.sugar),data=wq)

```

```{r echo=FALSE}

qplot(x=sqrt(chlorides),data=wq)

```


```{r echo=FALSE}

#trying to plot all the vaiables 
meltData <- melt(wq[2:13])

ggplot(meltData, aes(x = value)) + 
    facet_wrap(~variable,scales = "free_x") + 
    geom_histogram()

```

From Above plots we can see that residual.sugar and chlorides show right skewed
plot, while rest others are normally distributed.


# Univariate Analysis

### What is the structure of your dataset?
The main feature of interest is the quality of wine and how it will depend on
the different other variables like residual sugar,alcohol content, density etc.


Have not used any variation in the dataset. Didn't made any variable out of the
given variables.

# Bivariate Plots Section


```{r echo=FALSE}

ggplot(aes(x=quality,y=alcohol),data=wq)+
  geom_point(alpha=1/10,position=position_jitter(h=0),color='black',
             shape=21,fill=I('#F79420'))+
  xlim(0,15)

```


```{r echo=FALSE, Bivariate_Plots}

ggplot(aes(x=quality,y=log10(residual.sugar)),data=wq)+
  geom_point(alpha=1/10,position=position_jitter(h=0),color='black',
             shape=21,fill=I('#F79420'))+
  xlim(0,15)+
  ylim(0.0,0.8)
  

```

```{r echo=FALSE}

ggplot(aes(x=quality,y=sqrt(chlorides)),data=wq)+
  geom_point(alpha=1/10,position=position_jitter(h=0),color='black',
             shape=21,fill=I('#F79432'))+
  xlim(0,15)+
  ylim(0.2,0.4)

```

# Bivariate Analysis

```{r echo=FALSE}

with(wq,cor.test(alcohol,quality))

```


```{r echo=FALSE}

with(wq,cor.test(density,quality))

```

```{r echo=FALSE}

with(wq,cor.test(pH,quality))

```

```{r echo=FALSE}

with(wq,cor.test(total.sulfur.dioxide,quality))

```

```{r echo=FALSE}

with(wq,cor.test(citric.acid,quality))

```

```{r echo=FALSE}

with(wq,cor.test(fixed.acidity,quality))

```

```{r echo=FALSE}

with(wq,cor.test(residual.sugar,quality))

```

```{r echo=FALSE}

with(wq,cor.test(sulphates,quality))

```

Through above correlations we found out that the alcohol,sulphates, and citric
acid plays important part in deciding the quality of red wine.
Below making few plots to show the relation.

```{r echo=FALSE}

ggplot(aes(x=quality,y=sulphates),data=wq)+
  geom_point(alpha=1/10,position=position_jitter(h=0),color='black',
             shape=21,fill=I('#F79432'))

```

so we see that higher sulphate content wines are of better quality.


```{r echo=FALSE}

ggplot(aes(x=quality,y=citric.acid),data=wq)+
  geom_point(alpha=1/10,position=position_jitter(h=0),color='black',
             shape=21,fill=I('#F79432'))

```

Here we see that better quality wines have high citric acid content.

To test the relationships between various variables we will try to find out
the correlation between 2 variables and will try to find out if the relation 
is strong or not.



### Did you observe any interesting relationships between the other features \
#(not the main feature(s) of interest)?

```{r echo=FALSE,}

with(wq,cor.test(pH,fixed.acidity))

```
```{r echo=FALSE}

ggplot(aes(x=pH,y=fixed.acidity),data=wq)+
  geom_point()+
  geom_smooth()

```


As expected, the acidity is decreasing as the pH is increasing.


```{r echo=FALSE}

with(wq,cor.test(total.sulfur.dioxide,residual.sugar))

```

```{r echo=FALSE}

ggplot(aes(y=total.sulfur.dioxide,x=residual.sugar),data=wq)+
  geom_point()+
  geom_smooth()+
  xlim(0,4)

```

As expected the sulfur and sugar contents are not so very strongly correlated 
to each other. The graph doesnt show any big spikes in sulfur as sugar content 
is increasing.
So can't really say anything about this relationship.

### What was the strongest relationship you found?

The strongest relationship found is between pH to acidity and alcohol with 
respect to quality.

# Multivariate Plots Section


```{r echo=FALSE, Multivariate_Plots}
ggplot(aes(x=citric.acid,y=alcohol,color=factor(quality)),data=wq)+
  geom_point()+
  scale_color_brewer(type='div',
                     guide=guide_legend(title='Quality',reverse = TRUE))+
  ggtitle('Quality to citric acid & alcohol content')

```
As we can see that the best quality wines have good amount of alcohol ranging
more than 10 as well as some amount of citric acid. Lowe quality alcohol have 
more of citric acid than alcohol content.



```{r echo=FALSE}

ggplot(aes(x=pH,y=fixed.acidity,color=factor(quality),size=alcohol),data=wq)+
  geom_point()+
  scale_color_brewer(type='div',guide=guide_legend(title='Quality',
                                                   reverse = TRUE))+
  ggtitle('Quality to acidity, pH & alcohol content')

```
# Multivariate Analysis

The most strongest features for quality of red wines is alcohol content
rest others factors are secondary and doesnt matter much for quality.

The surprising factor is that its only alcohol which is directly influencing the
quality of wines. Other factors as pH fixed acidity do impact but few wines
still are of bad quality even though the alcohol pH and acidity are as they
shall be.


Have not created any Model for the dataset.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

qplot(x=quality,data=wq,binwidth=.1)

```

Most of the wines have quality of 5,6  rating. Highest quality wines are scarcely
available.


### Plot Two
```{r echo=FALSE, Plot_Two}


ggplot(aes(x=quality,y=alcohol),data=wq)+
  geom_point(alpha=1/10,position=position_jitter(h=0),color='black',
             shape=21,fill=I('#F79420'))+
  xlim(0,15)

```


### Description Two

The best quality leads to high alcohol content . The alcohol content is 
high even for medium quality wines.

### Plot Three
```{r echo=FALSE, Plot_Three}

ggplot(aes(x=citric.acid,y=alcohol,color=factor(quality)),data=wq)+
  geom_point()+
  scale_color_brewer(type='div',
                     guide=guide_legend(title='Quality',reverse = TRUE))+
  ggtitle('Quality to citric acid & alcohol content')


```

### Description Three

The good alcohol content the better quality but citric acid is also an important
factor to make good quality wines along with alcohol content. Its clearly visible
quality increases if the citric acid and alcoho both are mixed in right
proportion.
------

# Reflection

The exploration lead to the factor that alcohol and citric acid played well 
to get better quality wines. 
Main struggle was that of all the variables hardly 1-2 factors actually have
strong correlation with quality. So having intution that acidity, sulfur,
sugar content are strongly related to wine was totally negated.
Surprising was acidity didnt had much to do with quality.
We can make a data model for our dataset which i have not created.
